### 2020-07-13 appendix.tex
Updates from Overleaf

\section*{Appendix}

\section{Choices of Kernels}

We consider several string kernels below.

\begin{itemize}
    \item \textit{Spectrum Kernel.}
    \begin{align}
        k_\ell^{\text{Spec}}(\textbf{x}, \textbf{x}^\prime) =\left\langle\phi_{\ell}^{\mathrm{Spec}}(\mathbf{x}), \phi_{\ell}^{\mathrm{Spec}}\left(\mathbf{x}^{\prime}\right)\right\rangle = \phi_{\ell}^{\mathrm{Spec}}(\mathbf{x})^T \phi_{\ell}^{\mathrm{Spec}}\left(\mathbf{x}^{\prime}\right).
    \end{align}
     where $\mathbf{x}, \mathbf{x}^\prime$ are two RBS sequences in $\mathcal{D}$ over an alphabet $\Sigma$. Denote the number of letters in the alphabet as $|\Sigma|$. 
    $\phi_{\ell}^{\mathrm{spec}}(\mathbf{x})$ maps the sequence $\textbf{x}$ into a $|\Sigma|^\ell$ dimensional feature space, where each dimension is the count of the number of one of the $|\Sigma|^\ell$ possible strings $s$ of length $\ell$.  Let $\mathbf{X}, \mathbf{X}^\prime$ be two metrics which includes $n$ sequences, and $\Phi_d^{Spec}(\mathbf{X}) \in \mathbb{R}^{n \times |\Sigma|^{\ell}}$, the spectrum kernel over metrics is 
    \begin{align}
         K_\ell^{\text{Spec}}(\textbf{X}, \textbf{X}^\prime) = \Phi_{\ell}^{\mathrm{Spec}}(\mathbf{X}) \Phi_{\ell}^{\mathrm{Spec}}\left(\mathbf{X}^{\prime}\right)^T.
    \end{align}

    \item \textit{Sum of Spectrum Kernel,} considering weighted sum over different part of the string.

    \item \textit{Mixed Spectrum Kernel,} considering weighted sum over different substring length, with $\beta_d = \frac{2(\ell - d + 1)}{\ell(\ell+1)}$,
        \begin{align}
            k_\ell^{MixedSpec}(\mathbf{x}, \mathbf{x}^\prime) 
            = \sum_{d=1}^{\ell} \beta_d k_d^{Spec}(\mathbf{x}, \mathbf{x}^\prime)
        \end{align}
    \item \textit{Weighted Degree Kernel,} considering positional information. WD kernel counts the match of kmers at corresponding positions in two sequences.
    For sequences with fixed length $L$ and weighted degree kernel consider substrings starting at each position $l = 1, ..., L$, with $\beta_d = \frac{2(\ell - d + 1)}{\ell(\ell+1)}$, \\
    \begin{align}
        k_\ell^{WD}(\mathbf{x}, \mathbf{x}^\prime) 
        &= \sum_{d=1}^{\ell} beta_d \sum_{l=1}^{L-d+1} \gamma_l k_d^{Spec}(\mathbf{x}_{[l:l+d]}, \mathbf{x}_{[l:l+d]}^\prime)\\
        &= \sum_{d=1}^{\ell} beta_d \sum_{l=1}^{L-d+1} \gamma_l \phi_d^{Spec}(\mathbf{x}_{[l:l+d]})^T \phi_d^{Spec}(\mathbf{x}_{[l:l+d]}^\prime)\\
        &= \sum_{d=1}^{\ell} beta_d \sum_{l=1}^{L-d+1} \gamma_l \mathbb{I}(\mathbf{x}_{[l:l+d]} = \mathbf{x}_{[l:l+d]}^\prime),
    \end{align}
    where $\mathbb{I}(\text{true}) = 1$ and 0 otherwise.

    \item \textit{Weighted Degree Kernel With Shift.}
    \begin{align}
        k_\ell^{WDS}(\mathbf{x}, \mathbf{x}^\prime) 
        &= \sum_{d=1}^{\ell} \beta_d \sum_{l=1}^{L-d+1} \gamma_l \sum_{s = 0, s + l \leq L}^{S(l)} \delta_s
        \left(k_d^{Spec}(\mathbf{x}_{[l+s:l+s+d]}, \mathbf{x}_{[l:l+d]}^\prime) + (k_d^{Spec}(\mathbf{x}_{[l:l+d]}, \mathbf{x}_{[l+s:l+s+d]}^\prime)\right)\\
        &= \sum_{d=1}^{\ell} \beta_d \sum_{l=1}^{L-d+1} \gamma_l \sum_{s = 0, s + l \leq L}^{S(l)} \delta_s
        \left(\mathbb{I}(\mathbf{x}_{[l+s:l+s+d]} = \mathbf{x}_{[l:l+d]}^\prime) + (\mathbb{I}(\mathbf{x}_{[l:l+d]}= \mathbf{x}_{[l+s:l+s+d]}^\prime)\right),
    \end{align}
    where $\beta_d = \frac{2(\ell - d + 1)}{\ell(\ell+1)}, \delta_s = \frac{1}{2(s+1)}$, $\gamma_l$ is a weighting over the position in the
    sequence, which we set to 1. $S(l)$ determines the shift
    range at position $l$.
\end{itemize}

\textbf{From kernel to distance}:
$$d(\mathbf{x}, \mathbf{x}^\prime) = \sqrt{k(\mathbf{x}, \mathbf{x}) + k(\mathbf{x}^\prime, \mathbf{x}^\prime) - 2 k(\mathbf{x}, \mathbf{x}^\prime)} $$

\section{Design Choices}

\begin{itemize}
    \item Embedding.
    \item Kernel.
    \item Regression model.
    \item Bandit model.
\end{itemize}

\section{Intuition behind UCB and visualisation}

\begin{itemize}
    \item Exploitation and exploration explanation.
    \item Visualise coverage by clustering plot.
    \item Table for in-clustering mean and variance.
\end{itemize}

\section{Result analysis}

\begin{itemize}
    \item violinplot.
    \item regression performance plot, table.
    \item kernel matrix plot.
\end{itemize}

\section{Figures}