### 2020-08-08 appendix.tex
update kernel normalisation

\section*{Appendix}

\section{Choices of Kernels}

We consider several string kernels below.

\begin{itemize}
    \item \textit{Spectrum Kernel.}
    \begin{align}
        k_\ell^{\text{Spec}}(\textbf{x}, \textbf{x}^\prime) =\left\langle\phi_{\ell}^{\mathrm{Spec}}(\mathbf{x}), \phi_{\ell}^{\mathrm{Spec}}\left(\mathbf{x}^{\prime}\right)\right\rangle = \phi_{\ell}^{\mathrm{Spec}}(\mathbf{x})^T \phi_{\ell}^{\mathrm{Spec}}\left(\mathbf{x}^{\prime}\right).
    \end{align}
     where $\mathbf{x}, \mathbf{x}^\prime$ are two RBS sequences in $\mathcal{D}$ over an alphabet $\Sigma$. Denote the number of letters in the alphabet as $|\Sigma|$. 
    $\phi_{\ell}^{\mathrm{spec}}(\mathbf{x})$ maps the sequence $\textbf{x}$ into a $|\Sigma|^\ell$ dimensional feature space, where each dimension is the count of the number of one of the $|\Sigma|^\ell$ possible strings $s$ of length $\ell$.  Let $\mathbf{X}, \mathbf{X}^\prime$ be two metrics which includes $n$ sequences, and $\Phi_d^{Spec}(\mathbf{X}) \in \mathbb{R}^{n \times |\Sigma|^{\ell}}$, the spectrum kernel over metrics is 
    \begin{align}
         K_\ell^{\text{Spec}}(\textbf{X}, \textbf{X}^\prime) = \Phi_{\ell}^{\mathrm{Spec}}(\mathbf{X}) \Phi_{\ell}^{\mathrm{Spec}}\left(\mathbf{X}^{\prime}\right)^T.
    \end{align}

    \item \textit{Sum of Spectrum Kernel,} considering weighted sum over different part of the string.

    \item \textit{Mixed Spectrum Kernel,} considering weighted sum over different substring length, with $\beta_d = \frac{2(\ell - d + 1)}{\ell(\ell+1)}$,
        \begin{align}
            k_\ell^{MixedSpec}(\mathbf{x}, \mathbf{x}^\prime) 
            = \sum_{d=1}^{\ell} \beta_d k_d^{Spec}(\mathbf{x}, \mathbf{x}^\prime)
        \end{align}
    \item \textit{Weighted Degree Kernel,} considering positional information. WD kernel counts the match of kmers at corresponding positions in two sequences.
    For sequences with fixed length $L$ and weighted degree kernel consider substrings starting at each position $l = 1, ..., L$, with $\beta_d = \frac{2(\ell - d + 1)}{\ell(\ell+1)}$, \\
    \begin{align}
        k_\ell^{WD}(\mathbf{x}, \mathbf{x}^\prime) 
        &= \sum_{d=1}^{\ell} \beta_d \sum_{l=1}^{L-d+1} \gamma_l k_d^{Spec}(\mathbf{x}_{[l:l+d]}, \mathbf{x}_{[l:l+d]}^\prime)\\
        &= \sum_{d=1}^{\ell} \beta_d \sum_{l=1}^{L-d+1} \gamma_l \phi_d^{Spec}(\mathbf{x}_{[l:l+d]})^T \phi_d^{Spec}(\mathbf{x}_{[l:l+d]}^\prime)\\
        &= \sum_{d=1}^{\ell} \beta_d \sum_{l=1}^{L-d+1} \gamma_l \mathbb{I}(\mathbf{x}_{[l:l+d]} = \mathbf{x}_{[l:l+d]}^\prime),
    \end{align}
    where $\mathbb{I}(\text{true}) = 1$ and 0 otherwise.

    \item \textit{Weighted Degree Kernel With Shift.}
    \begin{align}
        k_\ell^{WDS}(\mathbf{x}, \mathbf{x}^\prime) 
        &= \sum_{d=1}^{\ell} \beta_d \sum_{l=1}^{L-d+1} \gamma_l \sum_{s = 0, s + l \leq L}^{S(l)} \delta_s
        \left(k_d^{Spec}(\mathbf{x}_{[l+s:l+s+d]}, \mathbf{x}_{[l:l+d]}^\prime) + (k_d^{Spec}(\mathbf{x}_{[l:l+d]}, \mathbf{x}_{[l+s:l+s+d]}^\prime)\right)\\
        &= \sum_{d=1}^{\ell} \beta_d \sum_{l=1}^{L-d+1} \gamma_l \sum_{s = 0, s + l \leq L}^{S(l)} \delta_s
        \left(\mathbb{I}(\mathbf{x}_{[l+s:l+s+d]} = \mathbf{x}_{[l:l+d]}^\prime) + (\mathbb{I}(\mathbf{x}_{[l:l+d]}= \mathbf{x}_{[l+s:l+s+d]}^\prime)\right),
    \end{align}
    where $\beta_d = \frac{2(\ell - d + 1)}{\ell(\ell+1)}, \delta_s = \frac{1}{2(s+1)}$, $\gamma_l$ is a weighting over the position in the
    sequence, which we set to 1. $S(l)$ determines the shift
    range at position $l$.
\end{itemize}

\textbf{From kernel to distance}:
$$d(\mathbf{x}, \mathbf{x}^\prime) = \sqrt{k(\mathbf{x}, \mathbf{x}) + k(\mathbf{x}^\prime, \mathbf{x}^\prime) - 2 k(\mathbf{x}, \mathbf{x}^\prime)} $$

\subsection{Normalisation of Kernel}

As part of data pre-processing,
the range of all features should be normalised so that each feature contributes approximately proportionately to the predictive model. 
The kernel matrix is represented by the inner product of the underlying feature vectors, it need to be normalised before using in the downstream regression models. 
Up-scaling (down-scaling) features can be understood as down-scaling (up-scaling) regularizers such that they penalise the features less (more).

Here we consider two approaches for kernel normalisation: centering and unit norm. 
We will show how to convert the normalisation in terms of feature vectors to normalisation in terms of kernel matrices. 
As defined before, consider $\mathbf{x}, \mathbf{x}^\prime$ are two RBS sequences in $\mathcal{D}$ over an alphabet $\Sigma$.
We denote $\phi(\mathbf{x}_i)$ as a column feature vector of sequence $\mathbf{x}_i$, 
where a feature function $\phi: \mathbf{x} \rightarrow \mathbb{R}^d$. Assume there are totally $n$ sequences in the data $\mathbf{X}$ ($n'$ sequences in the data $\mathbf{X}'$). 
We illustrate centering and unit norm normalisation below.

\begin{itemize}
    \item Centering. 
    Define the mean vector $\bar{\Phi}(\mathbf{X}) = \frac{1}{n} \sum_{s = 1}^n \phi(\mathbf{x}_s) \in \mathbb{R}^d$, the centered feature vector $\phi^C(\mathbf{x}_i) \in \mathbb{R}^d$ of $\mathbf{x}_i$ is
    \begin{align}
        \phi^{C}(\mathbf{x}_i) = \phi(\mathbf{x}_i) - \bar{\Phi}(\mathbf{X}) = \phi(\mathbf{x}_i) - \frac{1}{n} \sum_{s = 1}^n \phi(\mathbf{x}_s).
    \end{align}
    The corresponding centering kernel value between $\mathbf{x}_i$ and $\mathbf{x}_j$ is then 
    \begin{align}
        k^C(\mathbf{x}_i, \mathbf{x}_j) &= <\phi^C(\mathbf{x}_i), \phi^C(\mathbf{x}_j)>\\
        &= \left( \phi(\mathbf{x}_i) - \frac{1}{n} \sum_{s = 1}^n \phi(\mathbf{x}_s)\right)^T \left( \phi(\mathbf{x}_j) - \frac{1}{n} \sum_{s' = 1}^n \phi(\mathbf{x}_{s'})\right)\\
        &= \phi(\mathbf{x}_i)^T \phi(\mathbf{x}_j) - \left( \frac{1}{n} \sum_{s = 1}^n \phi(\mathbf{x}_s)\right)^T \phi(\mathbf{x}_j) - \phi(\mathbf{x}_i)^T \left(\frac{1}{n} \sum_{s' = 1}^n \phi(\mathbf{x}_{s'})\right) + \left( \frac{1}{n} \sum_{s = 1}^n \phi(\mathbf{x}_s)\right)^T \left(\frac{1}{n'} \sum_{s' = 1}^{n'} \phi(\mathbf{x}_{s'})\right)\\
        &= k(\mathbf{x}_i, \mathbf{x}_j) - \frac{1}{n} \sum_{s=1}^n k(\mathbf{x}_s, \mathbf{x}_j) - \frac{1}{n'} \sum_{s'=1}^{n'} k(\mathbf{x}_i, \mathbf{x}_{s'}) + \frac{1}{n^2} \sum_{s = 1}^n \sum_{s'=1}^{n'} k(\mathbf{x}_s, \mathbf{x}_{s'})
    \end{align}

    \item Unit Norm. Define the ($l_2$) norm of a feature vector $||\phi(\mathbf{x})|| = \sqrt{\sum_{m = 1}^d \phi_d(\mathbf{x})^2} = \sqrt{k(\mathbf{x}, \mathbf{x})} \in \mathbb{R}^+$, then the unit norm feature vector $\phi^{UN}(\mathbf{x}_i) \in \mathbb{R}^d$ of $\mathbf{x}_i$ is 
    \begin{align}
        \phi^{UN}(\mathbf{x}_i) = \frac{\phi(\mathbf{x}_i)}{||\phi(\mathbf{x}_i)||}.
    \end{align}
    The corresponding unit norm kernel value between $\mathbf{x}_i$ and $\mathbf{x}_j$ is then 
    \begin{align}
        k^{UN}(\mathbf{x}_i, \mathbf{x}_j) &= <\frac{\phi(\mathbf{x}_i)}{||\phi(\mathbf{x}_i)||}, \frac{\phi(\mathbf{x}_j)}{||\phi(\mathbf{x}_j)||}>\\
        &= \frac{\phi(\mathbf{x}_i)^T \phi(\mathbf{x}_j)}{||\phi(\mathbf{x}_i)|| \times ||\phi(\mathbf{x}_j)||}\\
        &= \frac{k(\mathbf{x}_i, \mathbf{x}_j)}{\sqrt{k(\mathbf{x}_i, \mathbf{x}_i)  k(\mathbf{x}_j, \mathbf{x}_j)}}
    \end{align}

     \item Unit Variance. 
    After the centering and unit norm normalisation, the kernel matrix is unit variance as well. 
    In the following, we show transformations of the unit variance (with centering) normalisation.
    Define the variance vector ${Var}(\Phi(\mathbf{X})) = \frac{1}{n} \sum_{s=1}^n ||\phi(\mathbf{x}_s) - \bar{\Phi}(\mathbf{X})||^2 = \frac{1}{n} \sum_{s=1}^n ||\phi(\mathbf{x}_s) - \sum_{s'=1}^n \left(\phi(\mathbf{x}_s')\right)||^2 = \frac{1}{n} \sum_{s=1}^n  k^C(\mathbf{x}_s, \mathbf{x}_s)  \in \mathbb{R}$, the unit variance feature vector $\phi^{UV}(\mathbf{x}_i) \in \mathbb{R}^d$ of $\mathbf{x}_i$ is
    \begin{align}
        \phi^{UV}(\mathbf{x}_i) = \frac{\phi(\mathbf{x}_i)}{\sqrt{Var(\Phi(\mathbf{X}))}}.
    \end{align}
    The corresponding kernel representation is 
    \begin{align}
        k^{UV}(\mathbf{x}_i, \mathbf{x}_j) &= <\frac{\phi(\mathbf{x}_i)}{\sqrt{Var(\Phi(\mathbf{X}))}}, \frac{\phi(\mathbf{x}_j)}{\sqrt{Var(\Phi(\mathbf{X}'))}}>\\
        &= \frac{\phi(\mathbf{x}_i)^T \mathbf{x}_j}{\sqrt{Var(\Phi(\mathbf{X})) Var(\Phi(\mathbf{X'}))}}\\
        &= \frac{k(\mathbf{x}_i, \mathbf{x}_j)}{\sqrt{ \frac{1}{n} \sum_{s=1}^n  k^C(\mathbf{x}_s, \mathbf{x}_s)  \frac{1}{n} \sum_{s'=1}^{n'}  k^C(\mathbf{x}_{s'}, \mathbf{x}_{s'})}}
    \end{align}
    After centering and unit norm, $ \frac{1}{n} \sum_{s=1}^n  k^C(\mathbf{x}_s, \mathbf{x}_s) = k(\mathbf{x}_i, \mathbf{x}_i)$, which implies that after centering and unit norm, the kernel matrix is already unit variance normalised. 
\end{itemize}
For the Gaussian Process regression, we make of use of two kernel matrices: the kernel function between the training data itself, i.e. $K(\mathbf{X}_{train}, \mathbf{X}_{train})$; 
the kernel function taking the training data and testing data as inputs, i.e. $K(\mathbf{X}_{test}, \mathbf{X}_{train})$. 
%It is straightforward to normalise a square kernel which same input, i.e. $n = n'$ and $k(\mathbf{x}_i, \mathbf{x}_i),  k(\mathbf{x}_j, \mathbf{x}_j)$ taken from the diagonal of the matrix. 
%The second one (between train and test) is a little bit tricky. The different is not only that $n \neq n'$. 
We will state two ways of normalise those two kind of matrices:
\begin{itemize}
    \item Normalise training and testing data separately.
    This approach is preferred for most of the machine learning algorithms since it follows the rule we have no information about testing data while training.
    Then for centering, one should subtract the mean vector over the training data for both kinds of matrices.
    For unit norm normalisation, when one calculate $K^{UN}(\mathbf{X}_{test}, \mathbf{X}_{train})$, the two terms inside of square root: $k(\mathbf{x}_i, \mathbf{x}_i)$ is taken from $K(\mathbf{X}_{test}, \mathbf{X}_{test})[i,i]$, and $k(\mathbf{x}_j, \mathbf{x}_j)$ is taken from $K(\mathbf{X}_{train}, \mathbf{X}_{train})[j,j]$.

    \item Normalise training and testing data together, i.e. normalise $K(\mathbf{X}_{train+test}, \mathbf{X}_{train+test})$, then extra the parts we need from the normalised matrix. 
    This approach is suitable for the case one already know the whole testing features. 
    For centering, one should subtract the mean vector over the whole matrix $\Phi(\mathbf{X}_{train+test})$. 
    The unit norm normalisation is the same as the previous case. 
\end{itemize}

For our experiment, we fix the design space before training, i.e. the testing features are already known before testing. 
So we choose to normalise the kernel matrix over the training and testing data together,
by first applying centering and then unit norm normalisation.

\section{Design Choices}

\begin{itemize}
    \item Embedding.
    \item Kernel.
    \item Regression model.
    \item Bandit model.
\end{itemize}

\section{Intuition behind UCB and visualisation}

\begin{itemize}
    \item Exploitation and exploration explanation.
    \item Visualise coverage by clustering plot.
    \item Table for in-clustering mean and variance.
\end{itemize}

\section{Result analysis}

\begin{itemize}
    \item violinplot.
    \item regression performance plot, table.
    \item kernel matrix plot.
\end{itemize}

\section{Figures}