### 2021-04-29 paper/mixture_absGau.tex

\newpage

\section{Mixture of Absolute Gaussian}

The raw TIR scores in each round are positive and  have two modes. 
We assume they are mixture of two absolute Gaussian distributions. 
If we can fit the mixture of absolute Gaussian distributions, we can answer the following questions:
\begin{itemize}
    \item is the left peak decreasing in height with each round?
	\item is the left peak going left?
	\item Is the right peak increasing in height with each round?
	\item is the right peak going right?
\end{itemize}

The density of absolute Gaussian (also called \textit{folded normal distribution}) is 
\begin{align}
    f_{Y}\left(x ; \mu, \sigma^{2}\right)=\frac{1}{\sqrt{2 \pi \sigma^{2}}} e^{-\frac{(z-\mu)^{2}}{2 \sigma^{2}}}+\frac{1}{\sqrt{2 \pi \sigma^{2}}} e^{-\frac{(x+\mu)^{2}}{2 \sigma^{2}}}
\end{align}
for $x \geq 0$ and $0$ everywhere else. 
The mean and variance of the absolute Gaussian distribution is 
\begin{align}
    \mu_{Y} &=\sigma \sqrt{\frac{2}{\pi}} \exp \left(\frac{-\mu^{2}}{2 \sigma^{2}}\right)+\mu \operatorname{erf}\left(\frac{\mu}{\sqrt{2 \sigma^{2}}}\right);\\
    \sigma_{Y}^{2} &=\mu^{2}+\sigma^{2}-\mu_{Y}^{2},
\end{align}
where $\operatorname{erf}$ is the error function 
$\operatorname{erf} z=\frac{2}{\sqrt{\pi}} \int_{0}^{z} e^{-t^{2}} d t$.

We consider the model of the mixture of two absolute Gaussian distributions as