### 2022-04-08 paper/supplementary.tex
fix ACS format

%! TEX root=./main.tex
\title{Supplementary of Machine learning guided\\
batched design of a bacterial Ribosome Binding Site}
\maketitle

\section{Machine Learning Methods}

In this section, we describe the machine learning methods in details.

\subsection{Gaussian Process Regression}
\label{supp: GPR}

A \textit{Gaussian process} is a collection of random variables, any finite number of which have a joint Gaussian distribution. 
We define mean function $\mu(\mathbf{x})$  and covariance function $k(\mathbf{x}, \mathbf{x}^\prime)$ of a real process $f(\mathbf{x})$ as
\begin{align}
    \mu(\mathbf{x}) &= \mathbb{E}[f(\mathbf{x})]\\
    k(\mathbf{x}, \mathbf{x}^\prime) &= \mathbb{E}[(f(\mathbf{x}) - \mu(\mathbf{x}))(f(\mathbf{x}^\prime) - \mu(\mathbf{x}^\prime))].
\end{align}

A Gaussian process is specified by its mean function and covariance function as $f(\mathbf{x}) \sim \mathcal{G} \mathcal{P}\left(\mu(\mathbf{x}), k\left(\mathbf{x}, \mathbf{x}^{\prime}\right)\right)$.
We consider the case where the observations are noisy, i.e. $(X, \mathbf{y}) = \{(\mathbf{x}_i, y_i)| i = 1, \dots, n\}$, where label $y_i = f(\mathbf{x}_i) + \epsilon$ with $\epsilon \sim \mathcal{N}(0, \alpha^2)$. 
The Gaussian noise is independent identically distributed, and the covariance of the prior on the noisy observations is then $\operatorname{cov}\left(y_{p}, y_{q}\right)=k\left(\mathbf{x}_{p}, \mathbf{x}_{q}\right)+\alpha^{2} \delta_{p q}$,
where $\delta_{pq}$ is a Kronecker delta which is one if $p = q$ and zero otherwise.
It is equivalent to a diagonal matrix $\alpha^2 I$ added on the kernel matrix evaluated on the training points.

For $n_\ast$ test points $X_\ast$, we assume the prior over the functions values as a random Gaussian vector $\mathbf{f}_\ast \sim \mathcal{N}(\mathbf{0}, K(X_\ast, X_\ast))$.
Then the joint distribution of the observed target values and the function values at the test points under the prior as 
\begin{align}
    \left[\begin{array}{l}\mathbf{y} \\ \mathbf{f}_{*}\end{array}\right] \sim \mathcal{N}\left(
        \left[\begin{array}{cc}
        \bm{\mu}  \\
        \bm{\mu}_\ast
        \end{array}\right],
    \left[\begin{array}{cc}K(X, X)+\alpha^{2} I & K\left(X, X_{*}\right) \\ K\left(X_{*}, X\right) & K\left(X_{*}, X_{*}\right)\end{array}\right]\right)
\end{align}
where
$\bm{\mu}, \bm{\mu}_\ast$ are mean vectors for $\mathbf{y}, \mathbf{f}_{*}$ respectively,
$K(X, X_\ast)$ denotes the $n \times n_\ast$ covariance/Kernel matrix evaluated at all pairs of training and testing points, similarly for other kernel matrices.
Then the posterior of the test points (i.e. predictive distributions) is given by the conditional distribution $\mathbf{f}_\ast | X, \mathbf{y}, X_\ast \sim \mathcal{N}(\bar{\mathbf{f}}_\ast, cov(\mathbf{f}_\ast))$, where
\begin{align}
   \overline{\mathbf{f}}_{*} & \triangleq \mathbb{E}\left[\mathbf{f}_{*} \mid X, \mathbf{y}, X_{*}\right]=
   \bm{\mu}_\ast + K\left(X_{*}, X\right)\left[K(X, X)+\alpha^{2} I\right]^{-1} (\mathbf{y} - \bm{\mu})\\
   \label{Eq: predicted variance}
   \operatorname{cov}\left(\mathbf{f}_{*}\right) &=K\left(X_{*}, X_{*}\right)-K\left(X_{*}, X\right)\left[K(X, X)+\alpha^{2} I\right]^{-1} K\left(X, X_{*}\right) 
\end{align}
For noisy test targets $\mathbf{y}_\ast = f(\mathbf{x}_\ast) + \epsilon$ with $\epsilon \sim \mathcal{N}(0, \alpha^2)$, we can compute the predictive distribution by adding $\alpha^2 I$ to the variance term $cov(\mathbf{f}_\ast)$ in Eq. (\ref{Eq: predicted variance}),
i.e. $\overline{\mathbf{y}}_{*} = \overline{\mathbf{f}}_{*}$, $\operatorname{cov}\left(\mathbf{y}_{*}\right) = \operatorname{cov}\left(\mathbf{f}_{*}\right) + \alpha^2 I$.

% We now introduce the marginal likelihood (or evidence) $p(\mathbf{y}|X)$, is which the integral of the likelihood times the prior 
% \begin{align}
%     p(\mathbf{y} \mid X)=\int p(\mathbf{y} \mid \mathbf{f}, X) p(\mathbf{f} \mid X) d \mathbf{f}
% \end{align}

\subsection{Choices of Kernels}
\label{sec: Choices of Kernels}

The choice of covariance function is critical for the performance of Gaussian process regression. 
Except the weighted degree kernel with shift described in Section \ref{sec: method prediction with kernel}, we show two closely related string kernels tested in this study below.

\begin{itemize}
    \item \textit{Spectrum Kernel.}
    \begin{align}
        k_\ell^{\text{Spec}}(\mathbf{x}, \mathbf{x}^\prime) =\left\langle\phi_{\ell}^{\mathrm{Spec}}(\mathbf{x}), \phi_{\ell}^{\mathrm{Spec}}\left(\mathbf{x}^{\prime}\right)\right\rangle = \phi_{\ell}^{\mathrm{Spec}}(\mathbf{x})^T \phi_{\ell}^{\mathrm{Spec}}\left(\mathbf{x}^{\prime}\right).
    \end{align}
     where $\mathbf{x}, \mathbf{x}^\prime$ are two RBS sequences in $\mathcal{D}$ over an alphabet $\Sigma$. We denote the number of letters in the alphabet as $|\Sigma|$. 
     In our case, $\Sigma = \{A,C,G,T\}$ and $|\Sigma| = 4$.
    $\phi_{\ell}^{\mathrm{spec}}(\mathbf{x})$ maps the sequence $\mathbf{x}$ into a $|\Sigma|^\ell$ dimensional feature space, where each dimension is the count of the number of one of the $|\Sigma|^\ell$ possible strings $s$ of length $\ell$. 
    For example, if $\mathbf{x} = ACACAG$ and $\ell = 3$, then $\phi_{\ell}^{\mathrm{spec}}(\mathbf{x})$ maps $\mathbf{x}$ into $4^3$ dimensional feature space, where each dimension corresponds to $AAA, AAC, AAT, AAG, ACA, ..., TTT$. Among those, the dimension corresponds to $ACA, CAC, CAG$ are 2,1,1 and others are 0. 
    Let $X, X^\prime$ be two metrics which include $n$ sequences, and $\Phi_\ell^{Spec}(X) \in \mathbb{R}^{n \times |\Sigma|^{\ell}}$, then the spectrum kernel over metrics is 
    \begin{align}
         K_\ell^{\text{Spec}}(X, X^\prime) = \Phi_{\ell}^{\mathrm{Spec}}(X) \Phi_{\ell}^{\mathrm{Spec}}\left(X^{\prime}\right)^T.
    \end{align}

    % \item \textit{Sum of Spectrum Kernel,} considers weighted sum over different parts of the string.

    % \item \textit{Mixed Spectrum Kernel,} considers weighted sum over different substring length, with $\beta_d = \frac{2(\ell - d + 1)}{\ell(\ell+1)}$,
    %     \begin{align}
    %         k_\ell^{MixedSpec}(\mathbf{x}, \mathbf{x}^\prime) 
    %         = \sum_{d=1}^{\ell} \beta_d k_d^{Spec}(\mathbf{x}, \mathbf{x}^\prime)
    %     \end{align}
    \item \textit{Weighted Degree Kernel,} considers positional information. WD kernel counts the match of kmers at corresponding positions in two sequences.
    For sequences with fixed length $L$ and weighted degree kernel considers substrings starting at each position $l = 1, ..., L$, with $\beta_d = \frac{2(\ell - d + 1)}{\ell(\ell+1)}$, \\
    \begin{align}
        k_\ell^{WD}(\mathbf{x}, \mathbf{x}^\prime) 
        &= \sum_{d=1}^{\ell} \beta_d \sum_{l=1}^{L-d+1} \gamma_l k_d^{Spec}(\mathbf{x}_{[l:l+d]}, \mathbf{x}_{[l:l+d]}^\prime)\\
        &= \sum_{d=1}^{\ell} \beta_d \sum_{l=1}^{L-d+1} \gamma_l \phi_d^{Spec}(\mathbf{x}_{[l:l+d]})^T \phi_d^{Spec}(\mathbf{x}_{[l:l+d]}^\prime)\\
        &= \sum_{d=1}^{\ell} \beta_d \sum_{l=1}^{L-d+1} \gamma_l \mathbb{I}(\mathbf{x}_{[l:l+d]} = \mathbf{x}_{[l:l+d]}^\prime),
    \end{align}
    where $\mathbb{I}(\text{true}) = 1$ and 0 otherwise.

    % \item \textit{Weighted Degree Kernel With Shift.}
    % \begin{align}
    %     k_\ell^{WDS}(\mathbf{x}, \mathbf{x}^\prime) 
    %     &= \sum_{d=1}^{\ell} \beta_d \sum_{l=1}^{L-d+1} \gamma_l \sum_{s = 0, s + l \leq L}^{S(l)} \delta_s
    %     \left(k_d^{Spec}(\mathbf{x}_{[l+s:l+s+d]}, \mathbf{x}_{[l:l+d]}^\prime) + (k_d^{Spec}(\mathbf{x}_{[l:l+d]}, \mathbf{x}_{[l+s:l+s+d]}^\prime)\right)\\
    %     &= \sum_{d=1}^{\ell} \beta_d \sum_{l=1}^{L-d+1} \gamma_l \sum_{s = 0, s + l \leq L}^{S(l)} \delta_s
    %     \left(\mathbb{I}(\mathbf{x}_{[l+s:l+s+d]} = \mathbf{x}_{[l:l+d]}^\prime) + (\mathbb{I}(\mathbf{x}_{[l:l+d]}= \mathbf{x}_{[l+s:l+s+d]}^\prime)\right),
    % \end{align}
    % where $\beta_d = \frac{2(\ell - d + 1)}{\ell(\ell+1)}, \delta_s = \frac{1}{2(s+1)}$, $\gamma_l$ is a weighting over the position in the
    % sequence, where we choose to use a uniform weighting over the sequences, i.e. $\gamma_l = 1/L$. $S(l)$ determines the shift
    % range at position $l$.
\end{itemize}

% \textbf{From kernel to distance}:
% $$d(\mathbf{x}, \mathbf{x}^\prime) = \sqrt{k(\mathbf{x}, \mathbf{x}) + k(\mathbf{x}^\prime, \mathbf{x}^\prime) - 2 k(\mathbf{x}, \mathbf{x}^\prime)} $$

\subsubsection{Normalisation of Kernel}
\label{supp: Normalisation of Kernel}

As part of data pre-processing,
the range of all features should be normalised so that each feature contributes approximately proportionately to the predictive model. 
The kernel matrix is represented by the inner product of the underlying feature vectors, it needs to be normalised before being used in the downstream regression models. 
Up-scaling (down-scaling) features can be understood as down-scaling (up-scaling) regularizers such that they penalise the features less (more).

Here we consider two approaches for kernel normalisation: centering and unit norm. 
We will show how to convert the normalisation in terms of feature vectors to normalisation in terms of kernel matrices. 
As defined before, consider $\mathbf{x}, \mathbf{x}^\prime$ are two RBS sequences in $\mathcal{D}$ over an alphabet $\Sigma$.
We denote $\phi(\mathbf{x})$ as a column feature vector of sequence $\mathbf{x}$, 
where a feature function $\phi: \mathbf{x} \rightarrow \mathbb{R}^d$,
with $d$ as the dimension of features.
One example of $\phi$ can be $\phi_\ell^{spec}(\mathbf{x})$ in the spectrum kernel.
Recall the corresponding kernel can be defined as $k(\mathbf{x}, \mathbf{x}^\prime) = \phi(\mathbf{x})^T \phi(\mathbf{x}).$

Assume there is total of $n$ sequences in the data $X$ ($n'$ sequences in the data $X'$). 
We illustrate centering and unit norm normalisation below.

\begin{itemize}
    \item Centering. 
    Defining the mean vector as $\bar{\Phi}(X) = \frac{1}{n} \sum_{s = 1}^n \phi(\mathbf{x}_s) \in \mathbb{R}^d$, the centered feature vector $\phi^C(\mathbf{x}_i) \in \mathbb{R}^d$ of $\mathbf{x}_i$ is
    \begin{align}
        \phi^{C}(\mathbf{x}_i) = \phi(\mathbf{x}_i) - \bar{\Phi}(X) = \phi(\mathbf{x}_i) - \frac{1}{n'} \sum_{s = 1}^{n'} \phi(\mathbf{x}_s).
    \end{align}
    The corresponding centering kernel value between $\mathbf{x}_i$ and $\mathbf{x}_j$ is then 
    \begin{align}
        k^C(\mathbf{x}_i, \mathbf{x}_j) &= <\phi^C(\mathbf{x}_i), \phi^C(\mathbf{x}_j)>\\
        &= \left( \phi(\mathbf{x}_i) - \frac{1}{n} \sum_{s = 1}^n \phi(\mathbf{x}_s)\right)^T \left( \phi(\mathbf{x}_j) - \frac{1}{n'} \sum_{s' = 1}^{n'} \phi(\mathbf{x}_{s'})\right)\\
        &= \phi(\mathbf{x}_i)^T \phi(\mathbf{x}_j) - \left( \frac{1}{n} \sum_{s = 1}^n \phi(\mathbf{x}_s)\right)^T \phi(\mathbf{x}_j) - \phi(\mathbf{x}_i)^T \left(\frac{1}{n} \sum_{s' = 1}^{n'} \phi(\mathbf{x}_{s'})\right) + \left( \frac{1}{n} \sum_{s = 1}^n \phi(\mathbf{x}_s)\right)^T \left(\frac{1}{n'} \sum_{s' = 1}^{n'} \phi(\mathbf{x}_{s'})\right)\\
        &= k(\mathbf{x}_i, \mathbf{x}_j) - \frac{1}{n} \sum_{s=1}^n k(\mathbf{x}_s, \mathbf{x}_j) - \frac{1}{n'} \sum_{s'=1}^{n'} k(\mathbf{x}_i, \mathbf{x}_{s'}) + \frac{1}{n^2} \sum_{s = 1}^n \sum_{s'=1}^{n'} k(\mathbf{x}_s, \mathbf{x}_{s'})
    \end{align}

    \item Unit Norm. Define the ($l_2$) norm of a feature vector as $||\phi(\mathbf{x})|| = \sqrt{\sum_{m = 1}^d \phi_m(\mathbf{x})^2} = \sqrt{k(\mathbf{x}, \mathbf{x})} \in \mathbb{R}^+$, then the unit norm feature vector $\phi^{UN}(\mathbf{x}_i) \in \mathbb{R}^d$ of $\mathbf{x}_i$ is 
    \begin{align}
        \phi^{UN}(\mathbf{x}_i) = \frac{\phi(\mathbf{x}_i)}{||\phi(\mathbf{x}_i)||}.
    \end{align}
    The corresponding unit norm kernel value between $\mathbf{x}_i$ and $\mathbf{x}_j$ is then 
    \begin{align}
        k^{UN}(\mathbf{x}_i, \mathbf{x}_j) &= <\frac{\phi(\mathbf{x}_i)}{||\phi(\mathbf{x}_i)||}, \frac{\phi(\mathbf{x}_j)}{||\phi(\mathbf{x}_j)||}>\\
        &= \frac{\phi(\mathbf{x}_i)^T \phi(\mathbf{x}_j)}{||\phi(\mathbf{x}_i)|| \times ||\phi(\mathbf{x}_j)||}\\
        &= \frac{k(\mathbf{x}_i, \mathbf{x}_j)}{\sqrt{k(\mathbf{x}_i, \mathbf{x}_i)  k(\mathbf{x}_j, \mathbf{x}_j)}}
    \end{align}

     \item Unit Variance. 
    After the centering and unit norm normalisation, the kernel matrix is unit variance as well. 
    In the following, we show transformations of the unit variance (with centering) normalisation.
    Define the variance vector ${Var}(\Phi(X)) = \frac{1}{n} \sum_{s=1}^n ||\phi(\mathbf{x}_s) - \bar{\Phi}(X)||^2 = \frac{1}{n} \sum_{s=1}^n ||\phi(\mathbf{x}_s) - \sum_{s'=1}^n \left(\phi(\mathbf{x}_s')\right)||^2 = \frac{1}{n} \sum_{s=1}^n  k^C(\mathbf{x}_s, \mathbf{x}_s)  \in \mathbb{R}$, the unit variance feature vector $\phi^{UV}(\mathbf{x}_i) \in \mathbb{R}^d$ of $\mathbf{x}_i$ is
    \begin{align}
        \phi^{UV}(\mathbf{x}_i) = \frac{\phi(\mathbf{x}_i)}{\sqrt{Var(\Phi(X))}}.
    \end{align}
    The corresponding kernel representation is 
    \begin{align}
        k^{UV}(\mathbf{x}_i, \mathbf{x}_j) &= <\frac{\phi(\mathbf{x}_i)}{\sqrt{Var(\Phi(X))}}, \frac{\phi(\mathbf{x}_j)}{\sqrt{Var(\Phi(X'))}}>\\
        &= \frac{\phi(\mathbf{x}_i)^T \mathbf{x}_j}{\sqrt{Var(\Phi(X)) Var(\Phi(\mathbf{X'}))}}\\
        &= \frac{k(\mathbf{x}_i, \mathbf{x}_j)}{\sqrt{ \frac{1}{n} \sum_{s=1}^n  k^C(\mathbf{x}_s, \mathbf{x}_s)  \frac{1}{n} \sum_{s'=1}^{n'}  k^C(\mathbf{x}_{s'}, \mathbf{x}_{s'})}}
    \end{align}
    After centering and unit norm, $ \frac{1}{n} \sum_{s=1}^n  k^C(\mathbf{x}_s, \mathbf{x}_s) = k(\mathbf{x}_i, \mathbf{x}_i)$, which implies that after centering and unit norm, the kernel matrix is already unit variance normalised. 
\end{itemize}
For the Gaussian Process regression, we make of use of two kernel matrices: the kernel function between the training data itself, i.e. $K(X_{train}, X_{train})$; and
the kernel function taking the training data and testing data as inputs, i.e. $K(X_{test}, X_{train})$. 
%It is straightforward to normalise a square kernel which same input, i.e. $n = n'$ and $k(\mathbf{x}_i, \mathbf{x}_i),  k(\mathbf{x}_j, \mathbf{x}_j)$ taken from the diagonal of the matrix. 
%The second one (between train and test) is a little bit tricky. The different is not only that $n \neq n'$. 
We will state two ways of normalisation those two kind of matrices:
\begin{itemize}
    \item Normalise training and testing data separately.
    This approach is preferred for most of the machine learning algorithms since it follows the rule that we have no information about testing data while training.
    Then for centering, one should subtract the mean vector over the training data for both kinds of matrices.
    For unit norm normalisation, when one calculates $K^{UN}(X_{test}, X_{train})$, the two terms inside of square root: $k(\mathbf{x}_i, \mathbf{x}_i)$ is taken from $K(X_{test}, X_{test})[i,i]$, and $k(\mathbf{x}_j, \mathbf{x}_j)$ is taken from $K(X_{train}, X_{train})[j,j]$.

    \item Normalise training and testing data together, i.e. normalise $K(X_{train+test}, X_{train+test})$, then extra the parts we need from the normalised matrix. 
    This approach is suitable in a case where one already knows the whole of testing features. 
    For centering, one should subtract the mean vector over the whole matrix $\Phi(X_{train+test})$. 
    The unit norm normalisation is the same as in the previous case. 
\end{itemize}

For our experiment, we fix the design space before training, i.e. the testing features are already known before testing. 
So we choose to normalise the kernel matrix over the training and testing data together (for Round 2-3, normalise over all RBS sequence in the design space),
by first applying centering and then unit norm normalisation.

\subsection{Batch Recommendation}

We consider recommending sequences in batch and using Gaussian Process Batch Upper Confidence Bound (GP-BUCB) algorithm  \cite{desautels2014parallelizing}.
We show technical details of GP-BUCB in the following.
With batches of size $B$, the feedback mapping $fb[t] = \lfloor(t-1) / B\rfloor B$, i.e. 
\begin{align}
    \mathrm{fb}[t]=\left\{\begin{array}{cl}
    0 & : t \in\{1, \ldots, B\} \\
    B & : t \in\{B+1, \ldots, 2 B\} \\
    2 B & : t \in\{2 B+1, \ldots, 3 B\} \\
    & \vdots
    \end{array}\right.
\end{align}

A key property of Gaussian Process regression is that the predictive variance in Eq. (\ref{Eq: predicted variance}) only depends on observed points (i.e. features), but not on the labels of these observed points. 
So one can compute the posterior variance without actually observing the labels. 
The GP-BUCB policy is to select sequences that
\begin{align}
    \operatorname{argmax}_{\mathbf{x}_i \in \mathcal{D}} \left( \mu_{fb[t]}(\mathbf{x}_i) + \beta_t \sigma_{t-1}(\mathbf{x}_i)\right).
\end{align}
And only update $y_{t^{\prime}}=f\left(\boldsymbol{x}_{t^{\prime}}\right)+\varepsilon_{t^{\prime}} \text { for } t^{\prime} \in\{\mathrm{fb}[t]+1, \ldots, \mathrm{fb}[t+1]\}$ at the end of each batch ($\mathrm{fb}[t]<\mathrm{fb}[t+1]$). 
This is equivalent to sequential GP-UCB with \textit{hallucinated observations} $\boldsymbol{y}_{\mathrm{fb}[t]+1: t-1}=\left[\mu_{\mathrm{fb}[t]}\left(\boldsymbol{x}_{\mathrm{fb}[t]+1}\right), \ldots, \mu_{\mathrm{fb}[t]}\left(\boldsymbol{x}_{t-1}\right)\right]$, while the posterior variance decreases.

\subsection{Bandit-0 Design}
\label{sec: ML design pipeline}
% \mengyan{explain somewhere why we have a different pipeline for round 0.}
% \mengyan{add some analysis on literature data?
% to verify 1) our library is "better" 2) it's hard to do ML on available library. some analysis in my mind (for literature data): number of data points, number of replicates (I remembered there only one replicates, if not, then std), label histogram, coverage of RBS.
% And we need to address the label in literature data and ours is "different", in the way that our pipeline is different for bandit-0 design}

The design of Round-0 is based on the literature data \cite{jervis2018machine}.
We first normalise the raw TIR to values between 0 and 1. 
We applied the Gaussian Process Regression with noise parameter $\alpha = 1e-10$. 
We chose to use one of the basic string kernels, the \textit{spectrum kernel} \cite{leslie2001spectrum} to process the core 6bp and dot product kernel \cite{Rasmussen2004} (with one-hot embedding) to process the 7bp flanking sequences both upstream and downstream of the core sequence.
The design size is 60 with UCB parameter $\beta = 1$.

\subsection{Additional Results}

As stated in Section \ref{subsec: results LEARN}, we merely aimed to provide a valid ranking in Learn part due to the small amount of noisy data, which can be captured by the Spearman correlation coefficient. 
We show $R^2$ score for the reader's interest, which quantifies value-based prediction evaluation. The $R^2$ increases along with rounds increase.

\begin{table}[h!]
\caption{$R^2$ score for predictions in Figure \ref{fig: Scatterplot}.}
\centering
\begin{tabular}{|c|c|c|c|c|}
\hline
                  & \textbf{\begin{tabular}[c]{@{}c@{}}Train 0\\ Test 1\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Train 0-1\\ Test 2\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Train 0-2\\ Test 3\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Train 80\%\\ Test 20\%\end{tabular}} \\ \hline
\textbf{$R^2$}       & 0.067                                                             & 0.226                                                               & 0.213                                                               & 0.269                                                                   \\ \hline
\textbf{Spearman} & 0.269                                                             & 0.468                                                               & 0.546                                                               & 0.479                                                                   \\ \hline
\end{tabular}
\end{table}

\subsection{Exploration-exploitation visualisation}
\label{subsec: Exploration-exploitation visualisation}

In this section, we show our method of exploration-exploitation visualisation displayed in Figure \ref{fig: Swarmplot and Quantplot}C.
The UCB score of a sequence $\mathbf{x}_i$ at Round $t$ is defined as 
\begin{align*}
% \label{Eq: UCB}
    \mu_{t-1}(\mathbf{x}_i) + \beta_t \sigma_{t-1}(\mathbf{x}_i),
\end{align*}
where $\beta_t$ is a hyperparameter balancing the exploration and exploitation, 
$\mu_{t-1}(\mathbf{x}_i), \sigma_{t-1}(\mathbf{x}_i)$ are the predicted TIR and standard deviation (STD) at Round $t-1$ for the sequence $\mathbf{x}_i$.
The sequences with high predicted TIR address exploitation, while the sequences with high predicted STD address exploration. 
We show the predicted TIR (x-axis) v.s. predicted STD (y-axis) in Figure \ref{fig:exploitationvsexplorationdetail} as orange points.

To compare the predicted TIR and STD in the same dimension (basis), we project the (predicted TIR, predicted STD) pair to the diagonal line (the grey dashline) set between (0,1) and (1,0) coordinates.
The slope of the diagonal line is selected to weight the predicted TIR and STD equally. 
Each point on the diagonal then, corresponds to an orthogonal projection (known as \emph{vector projection}) of an orange point.
As a result, the higher the predicted TIR the closer the point is to the (1,0) coordinate, and the higher the predicted STD the closer the point is to the (0,1) coordinate.
Finally, we colour those projected points using the RdBu spectrum, where the points with high STD (exploration) are shown in red and those with high mean TIR (exploitation) are shown in blue.
Finally, these colours correspond to ones shown in Figure \mbox{\ref{fig: Swarmplot and Quantplot}C}.

% \mengyan{mengyan to re-write it}
% For a given round all the tested sequences are then projected along the diagonal line crossing (0,1) and (1,0) according to their predicted STD and mean TIR (the higher the predicted mean the closer the point is to the (1,0) coordinate and the higher the predicted STD the closer the point is to the (0,1) coordinate).
% Each point on the diagonal then, corresponds to an orthogonal projection (known as \emph{vector projection}) of an orange point.
% We do this to transform our predictions into a common basis (where the basis is the diagonal) and so by calculating where a given point falls on the diagonal we can show if its mean TIR or STD is higher.
% Finally, we colour those projected points using the RdBu spectrum where the points with high STD (exploration) are shown in red and those with high mean TIR (exploitation) are shown in blue.

\begin{figure}[!ht]
   \centering
    \includegraphics[scale=0.35]{plots/Supplementary/proj_scatter_abc1_FF_0.pdf}
    \includegraphics[scale=0.35]{plots/Supplementary/proj_scatter_abc1_FF_1.pdf}
    \includegraphics[scale=0.35]{plots/Supplementary/proj_scatter_abc1_FF_2.pdf}
   \caption{\textbf{Exploration-Exploitation scores illustration.} We show the predictions for each sequence from Round 1-3 in orange (with the same settings as Figure \mbox{\ref{fig: Scatterplot}}) and their corresponding projections to (0,1), (1,0) diagonal coloured in RdBu spectrum.
    We explain in details how the colours are generated in Section \ref{subsec: Exploration-exploitation visualisation} .
   }
   \label{fig:exploitationvsexplorationdetail}
\end{figure}

\subsection{Model evaluation on RBS calculator data}
\label{subsec: Model evaluation on RBS calculator data}

\hl{To test the reliability of our proposed model, we have generated TIR predictions for all 4096 RBS sequences using the RBS  calculator ver. 2.0 \mbox{\cite{Salis2009}} and evaluated our model using that data.
We normalised the TIR values by dividing them by the TIR of the benchmark sequence, and show the TIR ratio mapping between the RBS calculator and TIRs obtained experimentally for this work in Figure \mbox{\ref{fig:TIR ratio salis micro}}.
We observe that the RBS calculator values are not well aligned to the measured values, which underlines the need for development of new prediction and design methods. 
Among all 4096 RBS sequences, 38 RBS were predicted to have higher TIR than the benchmark sequence by the RBS calculator. We call those sequences the \textit{optimal set}, and the goal is to find as many sequences from the optimal set as permitted by the budget (450).

To do this, we have applied our approach to this data set - first we have chosen 90 sequences at random and after teaching the algorithm using the obtained numbers we have generated 90 further sequences to be tested (using UCB for recommendations) and did this for a total of 4 rounds.
We have repeated this experiment 100 times to increase our confidence in the results, by doing so we have found that our model can identify above 75\% of the optimal set in up to 2 rounds and 99\% of them in up to 3 rounds.
We reported the detailed statistics for the 100 runs in Table \mbox{\ref{table: Important statistics for the 100 synthetic experiments}}.
Figure \mbox{\ref{fig: Swarmplot and Quantplot Salis}} shows plot analogues to these in Figure \mbox{\ref{fig: Swarmplot and Quantplot}}  but for one of the runs using the RBS calculator data.
Our algorithm has successfully found most of the sequences in optimal set in the first and second rounds, thus the recommended TIRs have an apparent drop in the following rounds (as there are no good sequences left to recommend).
This might suggest that it might be valuable to add an additional stopping condition for the DBTL cycle - decreasing round performance.}

\begin{figure}
    \centering
    \includegraphics[scale = 0.6]{paper/plots/Supplementary/TIR_ratio_salis_micro.png}
    \caption{TIR ratio of RBS calculator ver. 2.0 \cite{Salis2009} plotted against TIR ratios obtained experimentally in this work. The red point represents the benchmark sequence.}
    \label{fig:TIR ratio salis micro}
\end{figure}

\begin{figure}[!ht]
    \centering
    \begin{subfigure}[b]{0.48\textwidth}
        \centering
        \caption{}
        \includegraphics[scale=0.35]{plots/Supplementary/swarmplot_salis.pdf}
    \end{subfigure}
    % \hfill
    \begin{subfigure}[b]{0.25\textwidth}
        \centering
        \caption{}
        \includegraphics[scale=0.35]{plots/Supplementary/quantplot_salis.pdf}
    \end{subfigure}
    % \hfill
    \begin{subfigure}[b]{0.25\textwidth}
        \centering
        \caption{}
        \includegraphics[scale=0.35]{plots/Supplementary/swarmplot_proj_salis.pdf}
    \end{subfigure}
    \begin{subfigure}[b]{0.48\textwidth}
        \centering
        \caption{}
        \includegraphics[scale=0.4]{plots/Supplementary/histogram_salis.pdf}
    \end{subfigure}
    \begin{subfigure}[b]{0.48\textwidth}
        \centering
        \caption{}
        \includegraphics[scale=0.42]{plots/Supplementary/tsneplot_salis.pdf}
    \end{subfigure}
    % \includegraphics[scale=0.3]{plots/Main_Paper/swarmplot.pdf}
    % \includegraphics[scale=0.3]{plots/Main_Paper/quantplot.pdf}
    % \includegraphics[scale=0.3]{plots/Main_Paper/swarmplot_proj.pdf}
    % \includegraphics[scale=0.3]{plots/Main_Paper/histogram.pdf}
    % \includegraphics[scale=0.4]{plots/Main_Paper/tsneplot.pdf}
    \caption{
    \textbf{Model evaluation on RBS calculator data.}
    \textbf{A)} Swarm plot showing the obtained TIRs divided into RBS groups.
    Random: randomly chosen sequences.
    Bandit-0/1/2/3 - Bandit algorithm generated results for Round 0, 1, 2 and 3 respectively.
    Dashed line is set to TIR of the benchmark sequence.
    \textbf{B)} Line plot showing TIR obtained in a given quantile (Q) of results divided into groups as in A)
    \textbf{C)} Exploitation v.s. Exploration for Bandit 0-3. Blue-hued points represent exploitation, those hued red represent exploration.
    \textbf{D)} Histogram with kernel density estimations (KDE) showing distributions of TIRs for Bandit groups.
    \textbf{E)} t-SNE plot showing the relative distances between sequences in our design spaces as calculated by our kernel function (weighted degree kernel with shift).
    The area of the circle corresponds to the TIR value.
    The TIR results in all subplots are shown normalised to the  benchmark sequence.}
    \label{fig: Swarmplot and Quantplot Salis}
\end{figure}

\begin{figure}[!ht]
    \centering
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \caption{}
        \includegraphics[scale=0.4]{plots/Supplementary/scatter_abc1_FF_0_salis.pdf}
    \end{subfigure}
    % \hfill
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \caption{}
        \includegraphics[scale=0.4]{plots/Supplementary/scatter_abc1_FF_1_salis.pdf}
    \end{subfigure}
    % \hfill
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \caption{}
        \includegraphics[scale=0.4]{plots/Supplementary/scatter_abc1_FF_2_salis.pdf}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \caption{}
        \includegraphics[scale=0.4]{plots/Supplementary/scatter_abc1_FF_3_salis.pdf}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \caption{}
        \includegraphics[scale=0.4]{plots/Supplementary/salis_scatter_abc1_FF_4.pdf}
    \end{subfigure}
    \caption{\textbf{Performance of the prediction algorithm (no kernel normalisation) on the RBS calculator data}. The scatter plots A-D show the performance of our prediction algorithm calculated after each round using the RBS calculator data.
    E shows the performance on all collected data.
    Note that the TIR values are normalised according to the standardisation described in section \ref{sec: method data pre-procesing}, which is different from the TIR ratio reported in Figure \ref{fig: Swarmplot and Quantplot}.
    X-axis and y-axis are the TIR predicted by the RBS calculator and the TIR predicted by our model respectively.
    The Spearman correlation coefficient (with corresponding p-value; calculated using test data only) are provided in each plot's title on test data only.
    The p-value here is for the null hypothesis stating that two sets of data are uncorrelated.
    }
    \label{fig: Scatterplot_Salis}
\end{figure}

\begin{table}[h!]
\caption{\textbf{Important statistics for the 100 synthetic experiments run using the RBS calculator data.}
Each round contains 90 samples, where Round 0 is uniformly random sampled from the whole space, and Rounds 1-4 follow the bandits algorithm. 
In column \textit{Percent of the Optimal Set Found} and \textit{Standard Deviation}, we reported the average percent of the \textit{optimal set} found in each round and the standard deviation of this value averaged over the 100 experiments. 
In the last column, we reported the average percent of the (\textit{optimal set}) found in total after each round. 
We can observe that our algorithm can find the majority of the (\textit{optimal set}) sequences within three rounds.
}
\centering
\begin{tabular}{|c|c|c|c|}
\hline
                  \textbf{Round} & \textbf{\begin{tabular}[c]{@{}c@{}}Percent of\\ the Optimal Set Found\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Standard\\ Deviation\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Total Percent of\\ the Optimal Set Found\end{tabular}} \\ \hline
 \textbf{0}  & 0.022  & 0.026 & 0.022  \\ \hline
  \textbf{1}  & 0.744  & 0.085 & 0.766  \\ \hline
   \textbf{2}  & 0.222  & 0.077 & 0.988  \\ \hline
    \textbf{3}  & 0.012  & 0.026 & 1.000  \\ \hline
     \textbf{4}  & 0.000  & 0.000 & 1.000  \\ \hline

\end{tabular}
\label{table: Important statistics for the 100 synthetic experiments}
\end{table}

\newpage

\section{Supplementary Plots}

\begin{figure}[!ht]
    \centering
    \includegraphics[scale=0.4]{plots/Supplementary/core_vs_noncore.pdf}
    \caption{\textbf{Comparison of base change impact on TIR in core versus non-core region.} 
    The core region is highlighted in light green and the lines are rolling averages (in this case mean of the 5 points centred over the indicated position) for each base.
    The top dotted line shows the TIR for the benchmark sequence, where dots represent a change at a given position to a given base, which is colour coded.
    Since only changes from the original base at each position are shown, the dotted lines start and end at different positions depending on how many changes to the respective base are at positions adjacent to the indicated one.
    The value of Welch's t-test between the mean TIR in core and non-core groups is -4.8780 with p-value $<$ 0.0001 and 34 degrees of Freedom.}
    \label{fig:core_vs_noncore}
\end{figure}

\begin{figure}[!ht]
    \centering
    \includegraphics[scale = 0.4]{plots/Supplementary/swarmplot_raw.pdf}
    \caption{
    \textbf{Raw TIRs of RBS groups examined in this study.}
    Raw TIR is calculated as a derivative of GFP fluoresence divided by OD600 of culture over 4h counting from the start of log phase of growth. 
    BPS-NC: base-by-base changes in the non-core region. 
    BPS-C: base-by-base changes in the core region. 
    UNI: Randomly generated sequences with uniform distribution. 
    PPM: Randomly generated sequences with distribution following the PPM for all natural RBS in \emph{E. coli}. 
    Bandit-0/1/2/3 - Bandit algorithm generated results for Round 0, 1, 2 and 3 respectively.
    SD - Shine-Dalgarno sequence.
    BN - benchmark sequences for all plates. 
    }
    \label{fig:rawswarmplots.}
\end{figure}

\begin{figure}[!ht]
   \centering
    \includegraphics[scale=0.5]{plots/Supplementary/tsneplot_hamming.pdf}
   \caption{\textbf{t-SNE plot based on Hamming distances.} This t-SNE plot has been drawn based on Hamming distances between sequences. 
    }
   \label{fig: TSNE hamming}
\end{figure}

\begin{figure}[!ht]
    \centering
    \includegraphics[scale = 0.35]{plots/Supplementary/scatter_bc1_TTF_0.pdf}
    \includegraphics[scale = 0.35]{plots/Supplementary/scatter_abc1_TT_1.pdf}
    \includegraphics[scale = 0.35]{plots/Supplementary/scatter_abc1_TT_2.pdf}
    % \includegraphics[scale = 0.4]{plots/Supplementary/scatter_abc1_TT_3.pdf}
    \caption{\textbf{Predictions in design pipeline in Round 1-3.} For Round 1, the kernel matrix is normalised based on known RBS sequences in Round 0.
    For Round 2-3, the kernel matrix is normalised based on all RBS sequences in the design space.
    Compared to the results shown in Figure \mbox{\ref{fig: Scatterplot}}, here we have applied kernel normalisation described in Section \mbox{\ref{supp: Normalisation of Kernel}}. Although the precise value prediction looks different from Figure \mbox{\ref{fig: Scatterplot}}, the rankings provided in two figures are similar, which can be reflected by the similar Spearman correlation coefficients.
    }
    \label{fig:scatter abc1 TT.}
\end{figure}

% \begin{figure}[!ht]
%     \centering
%     \includegraphics[scale = 0.3]{plots/Supplementary/Data_Similarity_and_Prediction_Sorted_by_TIR_Label.pdf}
%     \includegraphics[scale = 0.3]{plots/Supplementary/Data_Similarity_and_Prediction_Sorted_by_RBS_Similarity.pdf}
%     \caption{Kernel Heatmap and Predictions. Sequences are grouped as Consensus (1-2), BPS-C (3-20), BPS-NC (21-61), UNI (62-90), PPM (91-118), Bandit-0 (119-177), Bandit-1 (178-265). Inside of each group, sequences are clustered and sorted in terms of TIR labels (left) or RBS similarity (right). The first row shows the similarity measured by weighted degree kernel with shift, the second shows the predicted mean and uncertainty (1.95 standard deviation).}
% \end{figure}

\begin{figure}[!ht]
   \centering
   \includegraphics[width=0.5\textwidth]{plots/Supplementary/All_logo.pdf}
   \caption{\textbf{Sequence logo for all tested sequences.} Compared to Figure \ref{fig:Library characteristics}A there are no significant biases at each position.}
   \label{fig: All_logo}
\end{figure}

\begin{figure}[!ht]
    \centering
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \includegraphics[scale=0.25]{plots/Supplementary/Structure_GGGGGG.pdf}
        \label{fig:GGGGGG}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \includegraphics[scale=0.25]{plots/Supplementary/Structure_GGGGGC.pdf}
        \label{fig:GGGGGC}
    \end{subfigure}
    \vskip\baselineskip
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \includegraphics[scale=0.25]{plots/Supplementary/Structure_CGGGAC.pdf}
        \label{fig:CGGGAC}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \includegraphics[scale=0.25]{plots/Supplementary/Structure_CACGTT.pdf}
        \label{fig:CACGTT}
    \end{subfigure}
       \caption{\textbf{Folding predictionds for four different RBS.} Here we show the predicted, energetically most favourable structures for four of our RBS with the following cores: GGGGGG, GGGGGC, CGGGAC, CACGTT and their immediate upstream and downstream background sequence. There is no discerbile and consistent difference between strong ones (the first two) and weak ones (the last two).}
       \label{fig:structures}
\end{figure}

\begin{figure}[!ht]
   \centering
    \includegraphics[scale=0.4]{plots/Supplementary/SDhist.pdf}
   \caption{\textbf{Histogram of standard deviations.}
   For each RBS sequence, we tested 6 biological replicates. This plot shows the histogram of the STD of raw TIR values of the 6 replicates for each RBS sequence. }
   \label{fig: SDhist}
\end{figure}

\begin{figure}[!ht]
    \centering
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \caption{}
        \includegraphics[scale=0.4]{plots/Supplementary/SD.pdf}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \caption{}
        \includegraphics[scale=0.6]{plots/Supplementary/SDvsTIR.pdf}
    \end{subfigure}
    \caption{\textbf{Distribution of standard deviation of tested samples} A) Joint plot showing the distribution of relative error calculated as the average TIR from 6 replicates divided by the given samples standard deviation (coefficient of variation). B) Scatter plot showing the standard deviation for each RBS plotted against its averaged TIR. Standard deviation and average TIR is given in terms of raw number.}
    \label{fig:variation of biological replicates}
\end{figure}